NMRlipids Databank makes data-driven analysis of biomembrane properties accessible for all

### 2023-04-20 main.tex
Update

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[3]{Lara S. Bort}             %ORCID: ?
\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[5]{Fernando Favela-Rosales}
\author[6]{Tiago Mendes Ferreira}
\author[7,8]{Patrick F.J. Fuchs}
\author[9]{Rebeca Garcia-Fandino}
\author[10]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[22]{Norbert Ku{\v c}erka}
\author[13]{Patrik Kula}
\author[14]{Milla Kurki}
\author[10]{Alexander Kuzmin}
\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[19]{Thomas Piggot}
\author[20]{{\'A}ngel Pi{\~n}eiro}
\author[10]{Suman Samantray}
\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
\author[1,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
%\affil[2]{Affiliation, department, city, postcode, country}
\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[10]{no affiliation}
\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[17]{Department of Chemistry, University of Bergen, Norway}
\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
\affil[22]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}

@@

P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
R.G.-F. thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09). 
AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
F. F.-R. acknowledges Tecnol{\'o}gico Nacional de M{\'e}xico, Direcci{\'o}n General de Asuntos del Personal Acad{\'e}mico (DGAPA), Programa de Apoyo a Proyectos de Investigaci{\'o}n e Innovaci{\'o}n Tecnol{\'o}gica (PAPIIT) 101923, CONACyT Ciencia de Frontera 74884, for financial support and Miztli-Direcci{\'o}n de C{\'o}mputo y de Tecnolog{\'i}as de Informaci{\'o}n y Comunicaci{\'o}n (DGTIC) - Universidad Nacional Aut{\'o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing time allocation.
M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
M.K acknowledge CSC – IT Center for Science, Finland, for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for the financial support

@@

A.M.K implemented the {\it Databank layer} and NMRlipids Databank-API.
H.S.A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
L.S.B. implemented about 630 pre-existing lipid bilayer simulations into the databank and created corresponding mapping files.
P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
F.F.-R. set up and performed simulations with Slipids Force Field: ID 617, ID 207 and ID 82. 
T.M.F. performed and analysed the solid-state NMR experiments.
P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
B.K. contributed to the initial design and implementation of the databank and was involved with the writing the manuscript.
N.K. contributed experimental scattering data. % and SIMtoEXP algorithm.
P.K. Implemented the flip-flop analysis.
M.K. provided simulations with OPLS3e force field. 
A.K. implemented the PCA-based equilibration metric.
J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing.
R.N. implemented the code for the calculations of X-ray scattering form factors.
A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
O.H.S.O. Designed and supervised the work, implemented water permeation and diffusion analyses, and wrote the manuscript.

>>>>

\author[1]{Anne M. Kiirikki}         %ORCID: ?
\author[2]{Hanne S. Antila}          %ORCID: 0000-0002-2474-5053
\author[3]{Lara S. Bort}             %ORCID: ?
\author[4]{Pavel Buslaev}         %ORCID: 0000-0003-2031-4691
\author[5]{Fernando Favela-Rosales}
\author[6]{Tiago Mendes Ferreira}
\author[7,8]{Patrick F.J. Fuchs}
\author[9]{Rebeca Garcia-Fandino}
\author[10]{Ivan Gushchin}
%\author[2]{Matti Javanainen}      %ORCID: 0000-0003-4858-364X
\author[11,12]{Batuhan Kav}           %ORCID: 0000-0003-4990-373X
\author[22]{Norbert Ku{\v c}erka}
\author[13]{Patrik Kula}
\author[14]{Milla Kurki}
\author[10]{Alexander Kuzmin}
\author[15,16]{Jesper J. Madsen}  %ORCID: 0000-0003-1411-9080
\author[17,18]{Markus S. Miettinen}   %ORCID: 0000-0002-3999-4722
%\author[2]{...}                   %ORCID: ?
\author[1]{Ricky Nencini}
\author[19]{Thomas Piggot}
\author[20]{{\'A}ngel Pi{\~n}eiro}
\author[10]{Suman Samantray}
\author[9,20,21]{Fabi{\'a}n Suarez-Leston}
\author[1,23,*]{O. H. Samuli Ollila} %ORCID: 0000-0002- 8728-1006
%\author[1,2,+]{Christine Author}
%\author[2,+]{Derek Author}

\affil[1]{University of Helsinki, Institute of Biotechnology, Helsinki, Finland}
\affil[2]{Department of Theory \& Bio-Systems and Department of Biomaterials, Max Planck Institute of Colloids and Interfaces, Germany}
\affil[3]{University of Potsdam, Institute of Physics and Astronomy, Potsdam-Golm, 14476, Germany}
%\affil[2]{Affiliation, department, city, postcode, country}
\affil[4]{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, P.O. Box 35, Jyv{\"a}skyl{\"a}, 40014 , Finland}
\affil[5]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, Sombrerete, Zacatecas, 99102, M\'{e}xico}
\affil[6]{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg,  Halle (Saale), 06120, Germany}
\affil[7]{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), Paris, 75005, France}
\affil[8]{Universit{\'e} Paris Cit{\'e}, UFR Sciences du Vivant, Paris, 75013, France}
\affil[9]{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela,  Santiago de Compostela, E-15782, Spain}
\affil[10]{no affiliation}
\affil[11]{Institute of Biological Information Processing: Structural Biochemistry (IBI-7), Forschungszentrum Jülich, Jülich 52428, Germany}
\affil[12]{ariadne.ai GmbH (Germany), Häusserstra{\ss}e 3 Heidelberg 69115, Germany }
\affil[13]{Institute of Organic Chemistry and Biochemistry of the Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2,  Prague, CZ-16610, Czech Republic}
\affil[14]{School of Pharmacy, University of Eastern Finland, 70211 Kuopio, Finland}
\affil[15]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[16]{Department of Molecular Medicine, Morsani College of Medicine, University of South Florida, Tampa, Florida, 33612, United States of America}
\affil[17]{Department of Chemistry, University of Bergen, Norway}
\affil[18]{Computational Biology Unit, Department of Informatics, University of Bergen, Norway}
\affil[19]{Chemistry, University of Southampton, Highfield, Southampton, SO17 1BJ, United Kingdom}
\affil[20]{Department of Applied Physics, Faculty of Physics, University of Santiago de Compostela, Santiago de Compostela, E-15782, Spain}
\affil[21]{MD.USE Innovations S.L., Edificio Emprendia, 15782 Santiago de Compostela, Spain}
\affil[22]{Department of Physical Chemistry of Drugs and Faculty of Pharmacy, Comenius University Bratislava, 832 32 Bratislava, Slovakia}
\affil[23]{VTT Technical Research Centre of Finland, Espoo, Finland}

@@

P.B. was supported by the Academy of Finland (Grants 311031 and 342908).
R.G.-F, A.P. and F.S.-L. thank Centro de Supercomputación de Galicia for computational support.
R.G.-F. thanks Ministerio de Ciencia, Innovaci{\'o}n y Universidades for a ''Ram{\'o}n y Cajal'' contract (RYC-2016-20335), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (RTI2018-098795-A-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431F 2020/05, 02\_IN606D\_2022\_2667887 and Centro singular de investigaci{\'o}n de Galicia accreditation 2016-2019, ED431G/09). 
AP thanks Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00, PDC2022-133402-I00), Xunta de Galicia and the ERDF (ED431B 2022/36, 02\_IN606D\_2022\_2667887).
F. S.-L. thanks Axencia Galega de Innovaci{\'o}n for his predoctoral contract (02\_IN606D\_2022\_2667887), and Spanish Agencia Estatal de Investigaci{\'o}n (AEI) and the ERDF (PID2019-111327GB-I00).
F. F.-R. acknowledges Tecnol{\'o}gico Nacional de M{\'e}xico, Direcci{\'o}n General de Asuntos del Personal Acad{\'e}mico (DGAPA), Programa de Apoyo a Proyectos de Investigaci{\'o}n e Innovaci{\'o}n Tecnol{\'o}gica (PAPIIT) 101923, CONACyT Ciencia de Frontera 74884, for financial support and Miztli-Direcci{\'o}n de C{\'o}mputo y de Tecnolog{\'i}as de Informaci{\'o}n y Comunicaci{\'o}n (DGTIC) - Universidad Nacional Aut{\'o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing time allocation.
M.S.M. was supported by the Trond Mohn Foundation BFS2017TMT01.
O.H.S.O, A.M.K, and R.N. acknowledge CSC --- IT Center for Science for computational resources and Academy of Finland (grants 315596 and 319902) for financial support.
M.K acknowledge CSC – IT Center for Science, Finland, for computational resources and thanks Finnish Cultural Foundation and the UEF Doctoral Programme for the financial support
We acknowledge all the NMRlipids project contributors for making the NMRlipids databank development possible.

@@

A.M.K implemented the {\it Databank layer} and NMRlipids Databank-API.
H.S.A. contributed to the quality evaluation metrics based on NMR and SAXS, and was involved in testing the GUI and polishing the manuscript.
L.S.B. implemented about 630 pre-existing lipid bilayer simulations into the databank and created corresponding mapping files.
P.B. implemented the PCA-based equilibration metric, supervised the work of A.K., and participated in code refactoring.
F.F.-R. set up and performed simulations with Slipids Force Field: ID 617, ID 207 and ID 82. 
T.M.F. performed and analysed the solid-state NMR experiments.
P.F.J.F. contributed to changes in the buildH software so that the Databank can handle united atom simulation data.
R.G.-F was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
I.G. advised on implementation of the PCA-based equilibration metric and supervised the work of A.K.
B.K. contributed to the initial design and implementation of the databank and was involved with the writing the manuscript.
N.K. contributed experimental scattering data. % and SIMtoEXP algorithm.
P.K. Implemented the flip-flop analysis.
M.K. provided simulations with OPLS3e force field. 
A.K. implemented the PCA-based equilibration metric.
J.J.M. provided simulations with Amber force field, general discussion, and manuscript editing.
R.N. implemented the code for the calculations of X-ray scattering form factors.
T.P. Performed many of the all-atom and united-atom simulations and contributed to discussions on the project.
A.P. was responsible for the design, validation, supervision and funding of the back-end and front-end for the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
F.S.-L. developed the back-end and participated in the design and development and validation of the graphical user interface (GUI) (http://www.databank.nmrlipids.fi/).
O.H.S.O. Designed and supervised the work, implemented water permeation and diffusion analyses, and wrote the manuscript.
